---
title: "Supplement TSCPB: Teachers’ sociodemographic characteristics and professional background"
date: today
published-title: "last modified"
author:
    name: Pawel R. Kulawiak
    orcid: 0000-0001-5939-4380
    email: kulawiak@uni-potsdam.de
format:
  html:
    title-block-banner: true
    number-sections: true
    embed-resources: true
    page-layout: full
    toc: true
    toc-location: body
    code-fold: true
    html-table-processing: none
execute: 
  warning: false
  message: false
bibliography: packages.bib
appendix-style: none
notice: |
  @R-base, @R-tidyverse, @R-gt, @R-readxl, @R-gtExtras, @R-devtools, @R-knitr
---

# R packages

```{r}
library(tidyverse)
library(gt)
library(readxl)
library(gtExtras)
library(devtools)
library(knitr)
write_bib(file = "packages.bib")
```

# Data import and manipulation

```{r}
DATA <- read_xlsx("SURVEY_DATA.xlsx")

DATA <-
  DATA %>%
  select(-where( ~ all(is.na(.)))) %>%
  filter(`teilnahme[teilnahme]` == "Y")

DATA <-
  DATA[9:81] %>%
  mutate(NA_CHECK = rowSums(!is.na(.))) %>%
  filter(NA_CHECK > 0) %>%
  filter(AB7 == 0) %>% # exclude special education schools
  mutate(SED_INC = case_when(AB10 == 0 & AB11 == 0 ~ 0,
                             AB10 == 1 & AB11 == 0 ~ 1,
                             AB10 == 0 & AB11 == 1 ~ 2,
                             AB10 == 1 & is.na(AB11) ~ 1,
                             is.na(AB10) & AB11 == 1 ~ 2,
                             AB10 == 1 & AB11 == 1 ~ 3,
                             is.na(AB10) & is.na(AB11) ~ NA,
                             AB10 == 0 & is.na(AB11) ~ NA,
                             is.na(AB10) & AB11 == 0 ~ NA)) %>%
  rename_with(~ stringr::str_replace_all(., "\\[", "_") %>%
                 stringr::str_replace_all(., "\\]", ""))
```

# Custom functions

```{r}
describe <-
  function(var) {
    DATA %>%
      group_by({{var}}) %>%
      reframe(n = n()) %>%
      rename(V1 = names(.)[1]) %>% 
      mutate(percent = (n / sum(n))) %>%
      mutate(INDEX = n) %>%
      mutate(percent = (percent*100) %>% round(1)) %>%
      relocate(percent, .after = last_col()) %>%
      mutate(n = NULL)
  }

school_subject <-
  function(var) {
    DATA %>%
      group_by({{var}}) %>%
      reframe(n = n()) %>%
      rename(V1 = names(.)[1]) %>% 
      mutate(percent = (n / sum(n))) %>%
      mutate(INDEX = n) %>%
      mutate(percent = (percent*100) %>% round(1)) %>%
      relocate(percent, .after = last_col()) %>%
      mutate(n = NULL) %>%
      na.omit()
  }

```

# Official school statistics (German Federal Statistical Office)

::: {.callout-note}
## Statistisches Bundesamt: Statistischer Bericht (Allgemeinbildende Schulen)													

- Schuljahr 2022/2023													
- EVAS-Nummer 21111													
- Ergänzung zur Datenbank GENESIS-Online
- Erschienen am 29. September 2023
- korrigiert am 04. Dezember 2023 (Tab. 21111-16 und csv-21111-16 für Bremen und Deutschland)
- Data file:
- <https://www.destatis.de/DE/Themen/Gesellschaft-Umwelt/Bildung-Forschung-Kultur/Schulen/Publikationen/Downloads-Schulen/statistischer-bericht-allgemeinbildende-schulen-2110100237005.xlsx?__blob=publicationFile>
- <https://www.destatis.de/DE/Themen/Gesellschaft-Umwelt/Bildung-Forschung-Kultur/Schulen/Publikationen/_publikationen-innen-schulen-allgemeinbildende.html>
:::


```{r}
OSS <- # official school statistics
  read_xlsx("statistischer-bericht-allgemeinbildende-schulen-2110100237005.xlsx", sheet = "csv-21111-21")
```

```{r}
SEX_OSS <-
  OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe == "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Grundschulen" |
           Schulart == "Hauptschulen" |
           Schulart == "Schularten mit mehreren Bildungsgängen" |
           Schulart == "Realschulen" |
           Schulart == "Gymnasien" |
           Schulart == "Integrierte Gesamtschulen") %>%
  filter(Geschlecht != "Insgesamt") %>%
  group_by(Geschlecht) %>%
  reframe(n_sex = sum(Lehrkraefte_Anzahl)) %>%
  mutate(n_percent = (n_sex / sum(n_sex)) * 100)
```

```{r}
AGE_OSS <-
  OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe != "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Grundschulen" |
           Schulart == "Hauptschulen" |
           Schulart == "Schularten mit mehreren Bildungsgängen" |
           Schulart == "Realschulen" |
           Schulart == "Gymnasien" |
           Schulart == "Integrierte Gesamtschulen") %>%
  filter(Geschlecht == "Insgesamt") %>%
  mutate(Altersgruppe = case_when(
    Altersgruppe %in% c("unter 30", "30-35") ~ "<34",
    Altersgruppe %in% c("35-40", "40-45") ~ "35-44",
    Altersgruppe %in% c("45-50", "50-55") ~ "45-54",
    Altersgruppe %in% c("55-60", "60-65", "65 und älter") ~ ">54") %>% as_factor()) %>%
  group_by(Altersgruppe) %>%
  reframe(n_age = sum(Lehrkraefte_Anzahl)) %>%
  mutate(n_percent = (n_age / sum(n_age)) * 100)
```

# Table: Teachers' sociodemographic characteristics

```{r}
NAMES_1 <-
  c("26-35", "36-45", "46-55", ">55",
    "male", "female", "prefer not to say", "missing value")

INDEX_2 <-
  c(AGE_OSS$n_age, SEX_OSS$n_sex, 0, 0)

percent_2 <-
  c(AGE_OSS$n_percent, SEX_OSS$n_percent, 0, 0)

rbind(
  describe(AB14), # age
  describe(AB15)  # sex
  ) %>%
  mutate(V1 = NAMES_1) %>% 
  mutate(INDEX_2 = INDEX_2) %>%
  mutate(percent_2 = percent_2) %>%
  gt() %>%
  gt_plt_bar_pct(column = percent,
                 scaled = T,
                 labels = TRUE,
                 fill = "black",
                 width = 100,
                 height = 20,
                 label_cutoff = 0.5,
                 font_size = "12px") %>%
  gt_plt_bar_pct(column = percent_2,
                 scaled = T,
                 labels = TRUE,
                 fill = "black",
                 width = 100,
                 height = 20,
                 label_cutoff = 0.5,
                 font_size = "12px") %>% 
  cols_align("left") %>%
  cols_label(V1 = "",
             INDEX = md("***N***"),
             percent = md("***%***"),
             INDEX_2 = md("***N***"),
             percent_2 = md("***%***")) %>%
  tab_row_group(label = md("*sex*"), rows = 5:8) %>%
  tab_row_group(label = md("*age in years^2^*"), rows = 1:4) %>%
  tab_style(style = list(cell_text(indent = pct(10))), locations = cells_body(columns = V1)) %>% 
  tab_options(column_labels.background.color = "gray95",
              footnotes.background.color = "gray95") %>%
  tab_style(style = cell_fill(color = "gray95"), locations = cells_row_groups(groups = everything())) %>%
  tab_spanner(label = md("*teacher<br>survey*"), columns = c(INDEX, percent)) %>%
  tab_spanner(label = md("*official<br>school statistics^1^*"), columns = c(INDEX_2, percent_2)) %>%
  tab_footnote(md("^1^*German Federal Statistical Office (EVAS-Number 21111):<br>
                  Official school statistics for Lower Saxony (school year 2022/2023)<br>
                  ^2^Age brackets used in this table are based on the teacher survey<br>
                  and differ slightly from those used in official school statistics:<br>
                  <34, 35-44, 45-54, >54*")) %>%
  cols_align(align = "center", columns = starts_with("per")) %>%
  cols_align(align = "right", columns = starts_with("IND")) %>%
  cols_width(starts_with("V1") ~ pct(33)) %>%
  tab_options(table.align = "left")
```

# Table: Teachers' professional background

```{r}
NAMES <-
  c("1-4 elementary education", "5-10 lower secondary education", "11-13 upper secondary education",
    "no", "special education", "inclusive education", "special and inclusive", "missing value",
    "<1", "1-2", "3-5", "6-10", "11-20", "21-30", "31-40",
    "German", "math", "natural science", "physical education", "other subjects",
    paste0(5:1, c(" very positive", "", "", "", " very negative")), "missing value")

rbind(
  school_subject(AB6_1), # grade 1-4
  school_subject(AB6_2), # grade 5-10
  school_subject(AB6_3), # grade 11-13
  describe(SED_INC), # teacher training
  describe(AB13), # Work experience as a teacher
  school_subject(AB9_1), # German
  school_subject(AB9_2), # math
  school_subject(AB9_3), # natural science
  school_subject(AB9_4), # physical education
  school_subject(AB9_5), # other subjects
  describe((AB12 - 6) %>% abs()) # attitude towards inclusive education
) %>%
  mutate(V1 = NAMES) %>% 
  gt() %>%
  gt_plt_bar_pct(column = percent,
                 scaled = T,
                 labels = TRUE,
                 fill = "black",
                 width = 100,
                 height = 20,
                 label_cutoff = 0.5,
                 font_size = "12px") %>% 
  cols_align("left") %>%
  cols_label(V1 = "",
             INDEX = md("***N***"),
             percent = md("***%***")) %>%
  tab_row_group(label = md("*attitude towards inclusive education^1^*"), rows = 21:26) %>%
  tab_row_group(label = md("*teaching... (multiple answers possible)*"), rows = 16:20) %>%
  tab_row_group(label = md("*work experience as a teacher in years*"), rows = 9:15) %>%
  tab_row_group(label = md("*teacher training in special or inclusive education*"), rows = 4:8) %>%
  tab_row_group(label = md("*teaching in grades... (multiple answers possible)*"), rows = 1:3) %>%
  tab_style(style = list(cell_text(indent = pct(5))), locations = cells_body(columns = V1)) %>% 
  tab_options(column_labels.background.color = "gray95",
  footnotes.background.color = "gray95") %>%
  tab_style(style = cell_fill(color = "gray95"), locations = cells_row_groups(groups = everything())) %>%
  cols_align(align = "center", columns = starts_with("percent")) %>%
  cols_align(align = "right", columns = starts_with("INDEX")) %>%
  cols_width(starts_with("V1") ~ pct(66)) %>%
  tab_options(table.align = "left") %>%
  tab_options(table.align = "left") %>%
  tab_footnote(md('^1^*The scoring of the rating scale has been reversed.
                   Scoring used<br>in the teacher survey:
                   1 = "very positive" and 5 = "very negative"*'))
```

## Elementary school teachers in Lower Saxony (official school statistics)

```{r}
OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe == "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Grundschulen" |
           Schulart == "Hauptschulen" |
           Schulart == "Schularten mit mehreren Bildungsgängen" |
           Schulart == "Realschulen" |
           Schulart == "Gymnasien" |
           Schulart == "Integrierte Gesamtschulen") %>%
  filter(Geschlecht == "Insgesamt") %>%
  group_by(Schulart) %>%
  reframe(n = sum(Lehrkraefte_Anzahl)) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  gt() %>% 
  tab_footnote("Grundschulen = Elementary schools") %>%
  tab_options(table.align = "left")
```

## Sex (elementary education vs. secondary education)

```{r}
DATA %>%
  filter(AB6_1 == "Y") %>% 
  group_by(AB15) %>%
  reframe(n = n()) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  mutate(AB15 = case_when(AB15 == 1 ~ "male", AB15 == 2 ~ "female", AB15 == 4 ~  "prefer not to say")) %>%  
  gt() %>%
  tab_header("elementary education (teacher survey)") %>%
  tab_options(table.align = "left")

DATA %>%
  filter(AB6_2 == "Y" | AB6_3 == "Y") %>% 
  group_by(AB15) %>%
  reframe(n = n()) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  mutate(AB15 = case_when(AB15 == 1 ~ "male", AB15 == 2 ~ "female", AB15 == 4 ~  "prefer not to say")) %>%  
  gt() %>%
  tab_header("secondary education (teacher survey)") %>%
  tab_options(table.align = "left")
```

```{r}
OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe == "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Grundschulen") %>%
  filter(Geschlecht != "Insgesamt") %>%
  group_by(Geschlecht) %>%
  reframe(n = sum(Lehrkraefte_Anzahl)) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  gt() %>%
  tab_header("elementary education (official school statistics)") %>% 
  tab_footnote(md("männlich = male<br>weiblich = female")) %>%
  tab_options(table.align = "left")

OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe == "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Hauptschulen" |
           Schulart == "Schularten mit mehreren Bildungsgängen" |
           Schulart == "Realschulen" |
           Schulart == "Gymnasien" |
           Schulart == "Integrierte Gesamtschulen") %>%
  filter(Geschlecht != "Insgesamt") %>% 
  group_by(Schulart) %>%
  mutate(total_within_schulart = sum(Lehrkraefte_Anzahl)) %>% 
  ungroup() %>% 
  group_by(Schulart, Geschlecht) %>%
  reframe(n = sum(Lehrkraefte_Anzahl),
          total = total_within_schulart) %>%
  mutate(percent = ((n / total) * 100) %>% round(1)) %>%
  gt() %>%
  tab_header("secondary education (official school statistics)") %>% 
  tab_footnote(md("männlich = male<br>weiblich = female")) %>%
  tab_options(table.align = "left")

OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe == "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Hauptschulen" |
           Schulart == "Schularten mit mehreren Bildungsgängen" |
           Schulart == "Realschulen" |
           Schulart == "Gymnasien" |
           Schulart == "Integrierte Gesamtschulen") %>%
  filter(Geschlecht != "Insgesamt") %>% 
  mutate(Schulart = fct_collapse(Schulart, secondary =
                                   c("Hauptschulen",
                                     "Schularten mit mehreren Bildungsgängen",
                                     "Realschulen",
                                     "Gymnasien",
                                     "Integrierte Gesamtschulen"))) %>%
  group_by(Schulart, Geschlecht) %>%
  reframe(n = sum(Lehrkraefte_Anzahl)) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  gt() %>%
  tab_header("secondary education (official school statistics)") %>% 
  tab_footnote(md("männlich = male<br>weiblich = female")) %>%
  tab_options(table.align = "left")
```

## Age (elementary education vs. secondary education)

```{r}
DATA %>%
  filter(AB6_1 == "Y") %>% 
  group_by(AB14) %>%
  reframe(n = n()) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  mutate(AB14 = c("26-35", "36-45", "46-55", ">55")) %>% 
  gt() %>%
  tab_header("elementary education (teacher survey)") %>%
  tab_options(table.align = "left")

DATA %>%
  filter(AB6_2 == "Y" | AB6_3 == "Y") %>% 
  group_by(AB14) %>%
  reframe(n = n()) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  mutate(AB14 = c("26-35", "36-45", "46-55", ">55")) %>% 
  gt() %>%
  tab_header("secondary education (teacher survey)") %>%
  tab_options(table.align = "left")
```

```{r}
OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe != "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Grundschulen") %>%
  filter(Geschlecht == "Insgesamt") %>%
  mutate(Altersgruppe = case_when(
    Altersgruppe %in% c("unter 30", "30-35") ~ "<34",
    Altersgruppe %in% c("35-40", "40-45") ~ "35-44",
    Altersgruppe %in% c("45-50", "50-55") ~ "45-54",
    Altersgruppe %in% c("55-60", "60-65", "65 und älter") ~ ">54") %>% as_factor()) %>%
  group_by(Altersgruppe) %>%
  reframe(n = sum(Lehrkraefte_Anzahl)) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  gt() %>%
  tab_header("elementary education (official school statistics)") %>%
  tab_options(table.align = "left")

OSS %>%
  filter(Bundesland == "Niedersachsen") %>% 
  filter(Altersgruppe != "Insgesamt") %>%
  filter(Beschaeftigungsumfang == "Voll- und Teilzeitbeschäftigte") %>%
  filter(Schulart == "Hauptschulen" |
           Schulart == "Schularten mit mehreren Bildungsgängen" |
           Schulart == "Realschulen" |
           Schulart == "Gymnasien" |
           Schulart == "Integrierte Gesamtschulen") %>%
  filter(Geschlecht != "Insgesamt") %>% 
  mutate(Schulart = fct_collapse(Schulart, secondary =
                                   c("Hauptschulen",
                                     "Schularten mit mehreren Bildungsgängen",
                                     "Realschulen",
                                     "Gymnasien",
                                     "Integrierte Gesamtschulen"))) %>%
  
  mutate(Altersgruppe = case_when(
    Altersgruppe %in% c("unter 30", "30-35") ~ "<34",
    Altersgruppe %in% c("35-40", "40-45") ~ "35-44",
    Altersgruppe %in% c("45-50", "50-55") ~ "45-54",
    Altersgruppe %in% c("55-60", "60-65", "65 und älter") ~ ">54") %>% as_factor()) %>%
  group_by(Schulart, Altersgruppe) %>%
  reframe(n = sum(Lehrkraefte_Anzahl)) %>%
  mutate(percent = ((n / sum(n)) * 100) %>% round(1)) %>%
  gt() %>%
  tab_header("secondary education (official school statistics)") %>%
  tab_options(table.align = "left")
```

# R session info

```{r collapse=TRUE}
session_info()
```

# References 
